Measurement Bias from Address Aliasing

### 2016-01-09 paper.tex
Language: things->strings.

\section{Bias from Environment Size}
\label{sec:environment}
System environment variables contain things like the name of the logged in user, home directory, shell state, and various other settings.
As a source of bias, it is not the environment variables themselves that are important, but rather the effect their total \emph{size} has on the position and alignment of the stack.
As indicated in Figure \ref{fig:virtualmemory}, environment variables and program arguments are allocated in the stack section of virtual memory, close to the upper address 0x7fff'ffffffff\footnote{Modern processors do not actually use the full 64-bit space, only the low order 47 bits are used for addressing memory.} and before the first call frame.
Changing environment variables will therefore offset the location of the stack, and consequently all stack allocated variables.
After this offset, the stack is normally realigned to a 16 byte boundary, enforced by the compiler.
Within a span of 4096 bytes there are thus 256 possible initial stack addresses, each representing a different execution context with respect to address aliasing. 
%Note that there is no clear relationship between environment size and stack location with ASLR enabled.
%However, there will still be as many execution contexts with respect to aliasing (considering the stack only), making any occurrences of measurement bias indeed random.

>>>>

\section{Bias from Environment Size}
\label{sec:environment}
System environment variables contain strings like the name of the logged in user, home directory, shell state, and various other settings.
As a source of bias, it is not the environment variables themselves that are important, but rather the effect their total \emph{size} has on the position and alignment of the stack.
As indicated in Figure \ref{fig:virtualmemory}, environment variables and program arguments are allocated in the stack section of virtual memory, close to the upper address 0x7fff'ffffffff\footnote{Modern processors do not actually use the full 64-bit space, only the low order 47 bits are used for addressing memory.} and before the first call frame.
Changing environment variables will therefore offset the location of the stack, and consequently all stack allocated variables.
After this offset, the stack is normally realigned to a 16 byte boundary, enforced by the compiler.
Within a span of 4096 bytes there are thus 256 possible initial stack addresses, each representing a different execution context with respect to address aliasing. 
%Note that there is no clear relationship between environment size and stack location with ASLR enabled.
%However, there will still be as many execution contexts with respect to aliasing (considering the stack only), making any occurrences of measurement bias indeed random.
